---
title: "Tracking Chinese Foreign Investments, 2005-2018"
subtitle: "with xaringan"
author: "Peter Magee"
date: "2022/05/01"
output:
  xaringan::moon_reader
---

```{r echo=FALSE, include=FALSE}
library(tidyverse)
chinese_investment <- read_csv("aei_fdi.csv")
library(knitr)
```

#Tracking Chinese Foreign Investments, 2005-2018

- Given constantly shifting geopolitical climates I thought it might be interesting to see where a global power like China chooses to investment its money from 2005-2018.

- Taking data from the American Enterprise Institute, we'll look at the sectors & regions where China focuses its foreign investments.

- Chinese foreign investment trends reveal the broader geopolitical tensions between the USA and China. 

---
#Tracking China's investment dollars in USA, 2005 - 2018
.pull-center[
```{r warning=FALSE}
investment_order <- c("LOGISTICS", "CHEMICALS", "METALS", "HEALTH", "AGRICULTURE", "ENTERTAINMENT", "OTHER", "TOURISM", "TRANSPORT", "REAL ESTATE", "ENERGY", "TECHNOLOGY", "FINANCE")

USA_by_sector <- chinese_investment %>%
  filter(countryInvestment == "USA") %>%
  mutate(amountInvestment = parse_number(amountInvestment)) %>%
  ggplot(aes(factor(sectorInvestment, level = investment_order), amountInvestment)) +
  geom_col() +
  scale_y_continuous(
    labels = scales::label_dollar(scale = .000000001, largest_with_cents = 0)
  ) +
  labs(
    y = "Investment amounts (Billions, USD)",
    x = "Sector",
    title = "How China invested its money in USA, 2005-2018"
  )

graph_1 <- USA_by_sector + coord_flip()
```
]

---
class: center, middle
```{r dpi= 320, echo=FALSE, eval=TRUE, warning=FALSE, fig.width=7, fig.height=4}
graph_1
```
---

#What US companies have received the most Chinese money?
.pull-center[
```{r warning=FALSE, results = 'hide', message=FALSE, error=TRUE}
chinese_investment %>%
  filter(countryInvestment == "USA" & !is.na(nameEntityInvestment)) %>%
  group_by(nameEntityInvestment) %>%
  mutate(amountInvestment = parse_number(amountInvestment)) %>%
  summarise(n = n(), total = amountInvestment / n) %>%
  arrange(desc(n, total)) %>%
  mutate(US_company = str_trim(nameEntityInvestment))

companies <- tibble(
  "Company Name" = c("UNOCAL", "Blackstone", "Citi Group", "Morgan Stanley", "Ingram Micro"),
  "Total Investment (billions)" = c("$18", "$17.6", "$10.5", "$9.6", "$6.0"),
  "Sector" = c("Energy", "Finance", "Finance", "Finance", "Technology")
)

comp_table <- kable(companies,
  caption = "US companies receiving most Chinese money",
  align = "c"
)
```
]
---
class: middle
.pull-center[
```{r dpi= 320, echo=FALSE, eval=TRUE, warning=FALSE, fig.width=7, fig.height=4}
comp_table
```
]
---
#What do investments look like by region?
.pull-center[
```{r warning=FALSE}
annum_invest <- function(x) {
  chinese_investment %>%
    filter(yearTransaction == x) %>%
    mutate(amountInvestment = parse_number(amountInvestment)) %>%
    ggplot(aes(regionInvestment, amountInvestment)) +
    geom_col() +
    scale_y_continuous(
      labels = scales::label_dollar(scale = .000000001, largest_with_cents = 0)
    ) +
    labs(
      y = "Investment amounts (Billions, USD)",
      x = "Region"
    ) +
    scale_fill_manual() +
    coord_flip()
}
```
]
---
#Chinese investments by region, 2005
.pull-center[
```{r dpi= 320, echo=FALSE, eval=TRUE, warning=FALSE, fig.width=7, fig.height=4}
annum_invest("2005")
```
]
---
#Chinese investments by region, 2012
.pull-center[
```{r dpi= 320, echo=FALSE, eval=TRUE, warning=FALSE, fig.width=7, fig.height=4}
annum_invest("2012")
```
]
---
#Chinese investments by region, 2018
.pull-center[
```{r dpi= 320, echo=FALSE, eval=TRUE, warning=FALSE, fig.width=7, fig.height=4}
annum_invest("2018")
```
]
---
#Looking more closely at European investments
.pull-center[
```{r}
country_sector <- function(x, y) {
  chinese_investment %>%
    filter(regionInvestment == x | yearTransaction == y) %>%
    mutate(amountInvestment = parse_number(amountInvestment)) %>%
    ggplot(aes(amountInvestment, sectorInvestment)) +
    geom_col() +
    scale_x_continuous(
      labels = scales::label_dollar(scale = .000000001, largest_with_cents = 0)
    ) +
    labs(
      y = "Sector",
      x = "Investment amounts (Billions, USD)"
    )
}
```
]
---
#Investments in Europe by sector, 2005
.pull-center[
```{r dpi= 320, echo=FALSE, eval=TRUE, warning=FALSE, fig.width=7, fig.height=4}
country_sector("EUROPE", "2005")
```
]
---
#Investments in Europe by sector, 2012
.pull-center[
```{r dpi= 320, echo=FALSE, eval=TRUE, warning=FALSE, fig.width=7, fig.height=4}
country_sector("EUROPE", "2012")
```
]
---
#Investments in Europe by sector, 2018
.pull-center[
```{r dpi= 320, echo=FALSE, eval=TRUE, warning=FALSE, fig.width=7, fig.height=4}
country_sector("EUROPE", "2018")
```
]
---
#Thank you!!!
- Dr. Barrett
- Michelle Wheatley
- American University
- DC NASA Space Grant Consortium 

#Reference:
Chinese Global Foreign Investment via American Enterprise Institute. (2019, January 18). Tidy Tuesday. https://gist.github.com/abresler/0f82312cd43021887d2998c3fe99fabd
